Empirical Analysis of Performance Bottlenecks in Graph Neural Network Training with GPUs

### 2020-12-25 NEUROC/sec3-experiment-design.tex
sec3修改了问题部分,sec4修改了标题

    \item[Q3] \emph{Which consumes most of memory in GNN training?} (Section~\ref{sec:memory_usage_analysis})

          The limited memory capacity of a GPU prevents us from training GNNs on big graphs.
          %
          We measured the peak memory usage during GNN training under different graph scales, input feature dimensions, and average degrees.
          %
          Based on the results, we analyzed which was the most memory-consuming component in a GNN.
          %
          Reducing its memory usage will enable us to train GNNs on bigger graphs under the same memory capacity.

    \item[Q4] \emph{Can sampling techniques remove performance bottlenecks in GNN training? Does the use of sampling techniques in training affect the model accuracy?} (Section~\ref{sec:effects_of_sampling_techniques_on_performance})

          Theoretically, sampling techniques can significantly reduce the number of graph neurons that participate in the training of a batch.
          %
          Consequently, the processing time and memory usage per batch should also decrease.
          %
          The model accuracy of GNNs may also be affected.
          %
          To validate effectiveness of sampling techniques, we measured the training time, peak memory usage and test accuracy under different batch sizes.
          %
          If sampling techniques are effective, they will be the keys to conduct GNN training on very big graphs.
          %
          If they are not effective, we want to find out which impairs its efficiency.
          %

>>>>

    \item[Q3] \emph{Which consumes most of memory in GNN training and inference?} (Section~\ref{sec:memory_usage_analysis})

          The limited memory capacity of a GPU prevents us from training GNNs on big graphs.
          %
          We measured the peak memory usage during GNN training/inference under different graph scales, input feature dimensions, and average degrees.
          %
          Based on the results, we analyzed which was the most memory-consuming component in a GNN.
          %
          Reducing its memory usage will enable us to train GNNs on bigger graphs under the same memory capacity.

    \item[Q4] \emph{Can sampling techniques remove performance bottlenecks in GNN training? What is the performance bottleneck of sample-based inference? Does the use of sampling techniques in training affect the model accuracy? } (Section~\ref{sec:effects_of_sampling_techniques_on_performance})

          Theoretically, sampling techniques can significantly reduce the number of graph neurons that participate in the training of a batch.
          %
          Consequently, the processing time and memory usage per batch should also decrease.
          %
          The model accuracy of GNNs may also be affected.
          %
          To validate effectiveness of sampling techniques, we measured the training time, peak memory usage and test accuracy under different batch sizes.
          %
          If sampling techniques are effective, they will be the keys to conduct GNN training on very big graphs.
          %
          If they are not effective, we want to find out which impairs its efficiency.
          %
          % TODO: 补充Sample-based Inference的介绍

### 2020-12-25 NEUROC/sec4-experiments.tex
更改了[]的引用
sec3修改了问题部分,sec4修改了标题

We answer the four questions in Section~\ref{sec:experimental_scheme} one by one with experiments.
%
Without otherwise mentioned, the reported training time per epoch was the average wall-clock training time of 50 epochs, excluding abnormal epochs.
%
During the training of some epochs, there were extra profiling overheads from NVIDIA Nsight Systems and GC pauses from the Python interpreter that significantly increased the training time.
%
We denoted the 25\% and 75\% quantiles of the training time of 50 epochs as Q1 and Q3, respectively.
%
We regarded the epochs with the training time \emph{outside} the range of $[Q1 - 1.5 * (Q3-Q1), Q3 + 1.5 * (Q3-Q1)]$ as abnormal epochs.

@@

For GAT, we modified the number of heads $K$ and the dimension of each head $d_{head}$ in the GAT layer 0.
%
The dimension of $\MyVec{h}^1_x$ is determined as $dim(\boldsymbol{h}^1_x) = K d_{head}$.
%
Thus, the computational costs of the GAT layer 0 and the GAT layer 1 should increase linearly with $K$ and $d_{head}$ separately. 
%
\figurename~\ref{fig:exp_hyperparameter_on_vertex_edge_phase_time_gat} confirms the theoretical analysis.

For GaAN, it is also based on the multi-head mechanism.
%
Its time complexity should be affected by $dim(\boldsymbol{h}^1_x)$ ($d^0_{out} = d^1_{in} = dim(\boldsymbol{h}^1_x)$), $d_a$, $d_v$, $d_m$, and the number of heads $K$.
%
\figurename~\ref{fig:exp_hyperparameter_on_vertex_edge_phase_time_gaan} demonstrates that the training time increased linearly with the hyper-parameters, except for $dim(\boldsymbol{h}^1_x)$.
As $dim(\boldsymbol{h}^1_x)$ increased, the training time increased first slightly and then linearly.
We observed similar phenomena in GCN, GGNN, and GAT:
When the values of hyper-parameters were too low, GNN training could not make full use of the computing power of the GPU.
%
When the values of hyper-parameters became high enough, training time increased linearly, supporting the time complexity analysis.

@@

\paragraph{Summary of Memory Usage}
%
The \emph{high} memory expansion ratio severely restricted the data scalability of GNN training.
%
The memory usage mainly came from the intermediate results of the \emph{edge calculation stage}.
%
Fixing the number of vertices, the memory usage increased \emph{linearly} along with the number of edges.
%
Fixing the GNN structure and the hyper-parameters, increasing the dimension of input feature vectors could reduce the memory expansion ratio.
%
To reduce the memory usage of GNN training, optimizations should focus on reducing memory footprints of the edge calculation stage.

@@

The average degree of the sampled subgraph was \emph{much lower} than the average degree of the original graph, especially when the relative batch size is low.
%
Taking the neighbor sampler with the relative batch size of 6\% as an example, the average degree of the \texttt{amp} dataset was 31.1, but the average degree of the sampled subgraph was only 5.8.
%
For the cluster sampler, the average degree was 3.0.
%
\figurename~\ref{fig:exp_sampling_minibatch_degrees_distribution} compares the degree distribution of the sampled subgraphs with the original graph.
%
The slopes of the curves were similar, indicating that the sampled subgraphs still followed the power-law degree distribution.
%
However, the numbers of high-degree vertices were much less than the original graph, lowering the average degrees.
%
According to the experimental results in Section~\ref{sec:training_time_breakdown}, if the average degree became lower, the proportion of the training time spent on the vertex calculation stage would become higher, especially for GGNN.

@@

\paragraph{Summary of Sampling Techniques}

@@

Since the size of the sampled subgraph was much smaller than the original graph, the accuracy of the GNN models got by sampling techniques might be different from the full-batch training.
%
To find out how the batch size affected the test accuracy, we trained GNNs with different batch sizes.
%
\figurename~\ref{fig:exp_sampling_relative_batch_size_accuracy_graphsage} and \figurename~\ref{fig:exp_sampling_relative_batch_size_accuracy_cluster} compare the test accuracy achieved by sampling techniques with the test accuracy of the full-batch training.
%
For each combination of dataset and GNN, we chose the hyper-parameters that achieved the highest accuracy according to the experimental results in Section 4.1.

@@

The relationships between batch size and test accuracy were complex.
%
A larger batch size did not always bring higher accuracy.
%
For example, the accuracy of GaAN in \figurename~\ref{fig:exp_cluster_sampling_accuracy_on_amp} and GGNN in \figurename~\ref{fig:exp_cluster_sampling_accuracy_on_fli} decreased as the batch size increased.
%
A smaller batch size sometimes could achieve higher accuracy.
%
For example, GAT achieved a higher accuracy with 1\% relative batch size than the full-batch training in \figurename~\ref{fig:exp_cluster_sampling_accuracy_on_amc}.
%
Given a sampling method, we found that the optimal batch size highly depended on the dataset and the GNN algorithm.
%
Our observations were similar to [31].
%
How to automatically select a proper batch size is a topic worth further studying.

>>>>

We answered the four questions in Section~\ref{sec:experimental_scheme} one by one with experiments.
%
Without otherwise mentioned, the reported training time per epoch was the average wall-clock training time of 50 epochs, excluding abnormal epochs.
%
During the training of some epochs, there were extra profiling overheads from NVIDIA Nsight Systems and GC pauses from the Python interpreter that significantly increased the training time.
%
We denoted the 25\% and 75\% quantiles of the training time of 50 epochs as Q1 and Q3, respectively.
%
We regarded the epochs with the training time \emph{outside} the range of $[Q1 - 1.5 * (Q3-Q1), Q3 + 1.5 * (Q3-Q1)]$ as abnormal epochs.

@@

For GAT, we modified the number of heads $K$ and the dimension of each head $d_{head}$ in the GAT layer 0.
%
The dimension of $\MyVec{h}^1_x$ was determined as $dim(\boldsymbol{h}^1_x) = K d_{head}$.
%
Thus, the computational costs of the GAT layer 0 and the GAT layer 1 should increase linearly with $K$ and $d_{head}$ separately. 
%
\figurename~\ref{fig:exp_hyperparameter_on_vertex_edge_phase_time_gat} confirms the theoretical analysis.

For GaAN, it was also based on the multi-head mechanism.
%
Its time complexity should be affected by $dim(\boldsymbol{h}^1_x)$ ($d^0_{out} = d^1_{in} = dim(\boldsymbol{h}^1_x)$), $d_a$, $d_v$, $d_m$, and the number of heads $K$.
%
\figurename~\ref{fig:exp_hyperparameter_on_vertex_edge_phase_time_gaan} demonstrates that the training time increased linearly with the hyper-parameters, except for $dim(\boldsymbol{h}^1_x)$.
As $dim(\boldsymbol{h}^1_x)$ increased, the training time increased first slightly and then linearly.
We observed similar phenomena in GCN, GGNN, and GAT:
When the values of hyper-parameters were too low, GNN training could not make full use of the computing power of the GPU.
%
When the values of hyper-parameters became high enough, training time increased linearly, supporting the time complexity analysis.

@@

\paragraph{Summary of Memory Usage in Training}
%
The \emph{high} memory expansion ratio severely restricted the data scalability of GNN training.
%
The memory usage mainly came from the intermediate results of the \emph{edge calculation stage}.
%
Fixing the number of vertices, the memory usage increased \emph{linearly} along with the number of edges.
%
Fixing the GNN structure and the hyper-parameters, increasing the dimension of input feature vectors could reduce the memory expansion ratio.
%
To reduce the memory usage of GNN training, optimizations should focus on reducing memory footprints of the edge calculation stage.

@@

The average degree of the sampled subgraph was \emph{much lower} than the average degree of the original graph, especially when the relative batch size was low.
%
Taking the neighbor sampler with the relative batch size of 6\% as an example, the average degree of the \texttt{amp} dataset was 31.1, but the average degree of the sampled subgraph was only 5.8.
%
For the cluster sampler, the average degree was 3.0.
%
\figurename~\ref{fig:exp_sampling_minibatch_degrees_distribution} compares the degree distribution of the sampled subgraphs with the original graph.
%
The slopes of the curves were similar, indicating that the sampled subgraphs still followed the power-law degree distribution.
%
However, the numbers of high-degree vertices were much less than the original graph, lowering the average degrees.
%
According to the experimental results in Section~\ref{sec:training_time_breakdown}, if the average degree became lower, the proportion of the training time spent on the vertex calculation stage would become higher, especially for GGNN.

@@

\paragraph{Summary of Sampling Techniques in Training}

@@

Since the size of the sampled subgraph was much smaller than the original graph, the accuracy of the GNN models got by sampling techniques might be different from the full-batch training.
%
To find out how the batch size affected the test accuracy, we trained GNNs with different batch sizes.
%
\figurename~\ref{fig:exp_sampling_relative_batch_size_accuracy_graphsage} and \figurename~\ref{fig:exp_sampling_relative_batch_size_accuracy_cluster} compare the test accuracy achieved by sampling techniques with the test accuracy of the full-batch training.
%
For each combination of dataset and GNN, we chose the hyper-parameters that achieved the highest accuracy according to the experimental results in Section 4.1.4.

@@

The relationships between batch size and test accuracy were complex.
%
A larger batch size did not always bring higher accuracy.
%
For example, the accuracy of GaAN in \figurename~\ref{fig:exp_cluster_sampling_accuracy_on_amp} and GGNN in \figurename~\ref{fig:exp_cluster_sampling_accuracy_on_fli} decreased as the batch size increased.
%
A smaller batch size sometimes could achieve higher accuracy.
%
For example, GAT achieved a higher accuracy with 1\% relative batch size than the full-batch training in \figurename~\ref{fig:exp_cluster_sampling_accuracy_on_amc}.
%
Given a sampling method, we found that the optimal batch size highly depended on the dataset and the GNN algorithm.
%
Our observations were similar to \cite{zeng2020_graphsaint}.
%
How to automatically select a proper batch size is a topic worth further studying.
